\Title

### 2015-10-13 moho-inversion.tex
Separate the equations for regularized inversion
This makes it easier to tie into bott's method

\begin{equation}
    \phi(\mbf{p}) = \|\mbf{d}^o - \mbf{d}(\mbf{p})\|_2^2
    = [\mbf{d}^o - \mbf{d}(\mbf{p})]^T[\mbf{d}^o - \mbf{d}(\mbf{p})].
\end{equation}

\noindent However, non-linear inversions for the relief of an interface (like
the Moho) are ill-posed and required additional constraints in the form of
regularization \citep{silva_potential-field_2001}.
We will use the first-order Tikhonov regularization to impose smoothness on the
solution,

\begin{equation}
    \theta(\mbf{p}) = \|\mbf{R}\mbf{p}\|^2_2
    = \mbf{p}^T\mbf{R}^T\mbf{R}\mbf{p},
\end{equation}

@@

\noindent $\Gamma(\mbf{p})$ is non-linear with respect to $\mbf{p}$.
Thus, we can find the minimum of the goal function using gradient-based
iterative optimization
methods like the Gauss-Newton or Steepest Descent methods.
Such methods start from an initial estimate $\mbf{p}^0$ and repeatedly update
the estimate until a minimum of the goal function is reached.

The update at the $k$th iteration $\mbf{\Delta p} = \mbf{p}^{k+1} - \mbf{p}^k$
for Gauss-Newton method is

\begin{equation}
    \mbf{H}^k\mbf{\Delta p} = -\mbf{\nabla}\Gamma^k,
    \label{eq:gaussnewton}
\end{equation}

\noindent in which
$\mbf{\nabla}\Gamma^k$ is the gradient vector and
$\mbf{H}^k$ is the Hessian matrix of $\Gamma$.

The gradient is given by

\begin{equation}
    \mbf{\nabla}\Gamma^k =
    -2\mbf{A}^T[\mbf{d}^o - \mbf{d}(\mbf{p}^k)] +
    2\mu\mbf{R}^T\mbf{R}\mbf{p}^k,
    \label{eq:gradient}
\end{equation}

\noindent
The first term of \eq{gradient} is the gradient of the data-misfit
function $\phi$ and the second term is the gradient of the regularizing
function $\theta$.

The Hessian matrix is given by

\begin{equation}
    \mbf{H}^k = 2\mbf{A}^T\mbf{A} +
    2\mu\mbf{R}^T\mbf{R}.
    \label{eq:hessian}
\end{equation}

\noindent
The first term of \eq{hessian} is the Gauss-Newton approximation of the Hessian
matrix of $\phi$ and the second term is the Hessian of $\theta$.

The Steepest Descent method uses only the gradient direction to
update the initial estimate,

\begin{equation}
    \mbf{\Delta p} = -\mbf{\nabla}\Gamma^k.
\end{equation}

\noindent
Because of this, the Steepest Descent method has poor convergence when the
current solution is close to the minimum of the goal function
\citep{kelley_iterative_1987}.

Both methods require the computationally expensive evaluation of the Jacobian
matrix $\mbf{A}$.
In practice, the derivative in \eq{jacobian} is often calculated through a
first-order finite-difference approximation.
Thus, evaluating $\mbf{A}$ requires $2\times M \times N$ forward modeling
operations for each iteration of the gradient descent algorithm.
Forward modeling using tesseroids requires several trigonometric function
evaluations and numerical integration.
\todo{Cite forward modeling paper}
Both factors contribute significantly to the total computation time required to
evaluate $\mbf{A}$.

>>>>

\begin{equation}
    \phi(\mbf{p}) =
    [\mbf{d}^o - \mbf{d}(\mbf{p})]^T[\mbf{d}^o - \mbf{d}(\mbf{p})].
    \label{eq:data-misfit}
\end{equation}

Function $\phi(\mbf{p})$ is non-linear with respect to $\mbf{p}$.
Thus, we can find its minimum using gradient-based
iterative optimization
methods like the Gauss-Newton or Steepest Descent methods.
Such methods start from an initial estimate $\mbf{p}^0$ and repeatedly update
the estimate until a minimum is reached.

For the Gauss-Newton method,
the update at the $k$th iteration $\mbf{\Delta p} = \mbf{p}^{k+1} - \mbf{p}^k$
is the solution of the linear system

\begin{equation}
    \mbf{H}^k\mbf{\Delta p} = -\mbf{g}^k,
    \label{eq:gaussnewton}
\end{equation}

\noindent in which
$\mbf{g}^k$ and $\mbf{H}^k$ are, respectively,
the gradient vector and the Hessian matrix of $\phi(\mbf{p})$.

The Steepest Descent method uses only the gradient direction
to update the initial estimate

\begin{equation}
    \mbf{\Delta p} = -\mbf{g}^k.
    \label{eq:steepest}
\end{equation}

\noindent
Thus, it does not require the computation and storage of the Hessian matrix
nor the solution of linear systems.
However, the Steepest Descent method has poor convergence when the
current solution is close to the minimum of the goal function
\citep{kelley_iterative_1987}.

The gradient vector and the Gauss-Newton approximation of the Hessian matrix
of $\phi(\mbf{p})$ are, respectively,

\begin{equation}
    \mbf{g}^k = -2\mbf{A}^T[\mbf{d}^o - \mbf{d}(\mbf{p}^k)],
    \label{eq:gradient}
\end{equation}

\noindent
and

\begin{equation}
    \mbf{H}^k = 2\mbf{A}^T\mbf{A},
    \label{eq:hessian}
\end{equation}

The computation and storage of the dense matrix $\mbf{A}$ is time consuming
and is often times the bottleneck of an inversion algorithm.
In practice, the derivative in \eq{jacobian} is often calculated through a
first-order finite-difference approximation.
Thus, evaluating $\mbf{A}$ requires $2\times M \times N$ forward modeling
operations for each iteration of the gradient descent algorithm.
Forward modeling using tesseroids requires several trigonometric function
evaluations and numerical integration.
\todo{Cite forward modeling paper}
Both factors contribute significantly to the total computation time required to
evaluate $\mbf{A}$.

\subsection{Regularization}

Non-linear inversions for the relief of an interface (like
the Moho) are ill-posed and required additional constraints in the form of
regularization \citep{silva_potential-field_2001}.
A common approach is to use the first-order Tikhonov regularization to impose
smoothness on the solution,

\begin{equation}
    \theta(\mbf{p}) = \mbf{p}^T\mbf{R}^T\mbf{R}\mbf{p},
\end{equation}

@@

\noindent
The regularization parameter $\mu$ controls the balance between fitting the
observed data and obeying the smoothness constraint.

The goal function $\Gamma(\mbf{p})$ is also non-linear with respect to
$\mbf{p}$ and can be minimized using the Gauss-Newton or Steepest Descent
methods.
The gradient vector and Hessian matrix of the goal function are, respectively,

\begin{equation}
    \mbf{g}^k =
        -2\mbf{A}^T[\mbf{d}^o - \mbf{d}(\mbf{p}^k)] +
        2\mu\mbf{R}^T\mbf{R}\mbf{p}^k,
    \label{eq:gradient-regul}
\end{equation}

\noindent and

\begin{equation}
    \mbf{H}^k = 2\mbf{A}^T\mbf{A} + 2\mu\mbf{R}^T\mbf{R}.
    \label{eq:hessian-regul}
\end{equation}
